Distributed Nonlinear Feedback Design with Separable Control-Contraction Metrics

### 2016-12-02 R0V0.tex
minor change
First formulation of the ADMM paragraph using the Gramians

\paragraph{Notation} Let $\mathbb{S}$ be a totally ordered set and $s_1,s_2,s_3\in\mathbb{S}$. The notation $\mathbb{S}_{[s_1,s_2)}$ (resp. $\mathbb{S}_{\diamond s_3}$) stands for the set $\{s\in\mathbb{S}:s_1\leq s< s_2\}$ (resp. $\{s\in\mathbb{S}:s\diamond s_3\}$, where $\diamond$ is a comparison operator, i.e., $\diamond\in\{<,\geq,=,\ \text{etc}\}$). Let $n>0$ be any integer, the vector $e_i$ denotes the vector with zeros in all rows except the $i$-th where it is 1. A matrix $M\in\mathbb{R}^{n\times n}$ with zero elements except (possibly) those  $m_{ii},\ldots,m_{nn}$ on the diagonal is denoted as $\mathbin{\mathtt{diag}}(m_{ii},\ldots,m_{nn})$. The notation $M\succ 0$ (resp. $M\succeq 0$) stands for $M$ being positive (resp. semi)definite. For a matrix $M\succ0$, the notation $\mathbin{\mathtt{chol}}(M)$ stands for the \emph{upper-triangular Cholesky factor $U$} satisfying $A=U^\top U$.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control-contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x\neq0,\quad\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{align}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x<\ -2\lambda \delta_x^\top M\delta_x\label{eq:}
		\end{align}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
\mypara The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition to the  exponential convergence of solutions. Given a control-contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides stabilizing a feedback law of the form $\delta_u=K\delta_x$ for system \eqref{eq:general system:differential} defined for every $\delta_x\in\mathbb{R}^n$, where $K:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{n\times m}$.

@@

\section{Design of Decentralized Controllers}\label{sec:Design of Decentralized Controllers}

@@

\mypara The alternating direction method of multipliers (ADMM) provides a framework to compute the block-diagonal matrix $W$ to satisfy the MI~\eqref{eq:MI formulation} in a distributed fashion \cite{BoydParikhChuEtAl2010,MeissenLessardArcakEtAl2015}. To use this framework, define the sets $\mathbf{S}_{\succ0}^n=\{W\in\mathbb{R}^{n\times n}:W\succ0,W=W^\top\}$, $\mathbf{T}=\{W_1,\ldots,W_N:\exists Y\in\Xi,(\mathbin{\mathtt{diag}}(W_1,\ldots,W_N),Y)\text{ is a solution to \eqref{eq:MI formulation}}\}$. For a set $\mathbf{G}$ define the indicator/barrier function $H:\{\mathbf{G}\}\times \mathbb{R}^k\to \{0\}\cup\{\infty\}$ as
\begin{equation*}
	H(\mathbf{G},g)=\begin{cases}
									0,&g\in \mathbf{G}\\
									\infty,&\text{Otherwise.}
								\end{cases}
\end{equation*}

\mypara The MI~\eqref{eq:MI formulation} can be cast into the following optimization problem
\begin{equation*}
	\begin{array}{rl}
		\text{minimize}&\displaystyle H(\mathbf{T},(Z_1,\ldots,Z_N))+\sum_{i=1}^N H(\mathbf{S}_{\succ0}^{n_i},W_i)\\
		\text{subject to}&W_i-Z_i=0,\quad\forall i\in\mathbb{N}_{[1,N]}\;.
	\end{array}
\end{equation*}
The ADMM algorithm can be written as
\begin{enumerate}
	\item For each $i\in\mathbb{N}_{[1,N]}$, update the matrix $W_i$ as follows.
	\begin{equation*}
	\begin{array}{rcl}
		W_i^{k+1}&=&\mathbin{\mathtt{argmin}}_{W_i\in\mathbf{S}_{\succ0}^{n_i}}\left|\left|W_i-Z_i^k+\Lambda_i^k\right|\right|_F^2\;,
	\end{array}
\end{equation*}
	where $||\cdot||_F$ stands for the Frobenius norm;

	\item If the matrices $W_1,\ldots,W_N$ belong to $T$, then a solution has been found and the algorithm can be stopped. Otherwise, solve the global problem
	\begin{equation*}
		Z^{k+1}=\mathbin{\mathtt{argmin}}_{Z\in\mathbf{T}}\sum_{i=1}^N\left|\left|W_i^{k+1}-Z_i+\Lambda_i^k\right|\right|_F^2\;;
	\end{equation*}
	\item Update $\Lambda$ and return to step 1
	\begin{equation*}
		\Lambda_i^{k+1}=X_i^{k+1}-Z_i^{k+1}+\Lambda_i^k\;.
	\end{equation*}
\end{enumerate}
Note that steps 1 and 3 are computed in parellel while step 2 requires to gather all the values of the matrices $W_i$ and the values $\Lambda_i$.

>>>>

\paragraph{Notation} Let $\mathbb{S}$ be a totally ordered set and $s_1,s_2,s_3\in\mathbb{S}$. The notation $\mathbb{S}_{[s_1,s_2)}$ (resp. $\mathbb{S}_{\diamond s_3}$) stands for the set $\{s\in\mathbb{S}:s_1\leq s< s_2\}$ (resp. $\{s\in\mathbb{S}:s\diamond s_3\}$, where $\diamond$ is a comparison operator, i.e., $\diamond\in\{<,\geq,=,\ \text{etc}\}$). Let $n>0$ be any integer, the vector $e_i$ denotes the vector with zeros in all rows except the $i$-th where it is 1. A matrix $M\in\mathbb{R}^{n\times n}$ with zero elements except (possibly) those  $m_{ii},\ldots,m_{nn}$ on the diagonal is denoted as $\mathbin{\mathtt{diag}}(m_{ii},\ldots,m_{nn})$. The notation $M\succ 0$ (resp. $M\succeq 0$) stands for $M$ being positive (resp. semi)definite such a class of matrices is denoted as $\mathbf{S}_{\star0}^n=\{M\in\mathbb{R}^{n\times n}:M\star0,M=M^\top\}$, where $\star\in\{\succ,\succeq,\prec,\preceq\}$. For a matrix $M\succ0$, the notation $\mathbin{\mathtt{chol}}(M)$ stands for the \emph{upper-triangular Cholesky factor $U$} satisfying $A=U^\top U$.

@@

\begin{definition}[{\cite{Manchester2014a}}]\label{def:}
	A metric for system \eqref{eq:general system} is said to be a \emph{control-contraction metric for system \eqref{eq:general system}} if there exists a constant value $\lambda>0$ such that the condition
	\begin{subequations}\label{eq:Arstein-Sontag}
		\begin{equation}\label{eq:}
			\delta_x\neq0,\quad\delta_x^\top M(x)B(x)=0
		\end{equation}
		implies that the inequality
		\begin{equation}\label{eq:Artstein-Sontag:inequality}
			\delta_x^\top(\dot{M}+A^\top M+MA)\delta_x<\ -2\lambda \delta_x^\top M\delta_x
		\end{equation}
		holds, where $\dot{M}:=\partial_{f+Bu}M$.
	\end{subequations}
\end{definition}
\mypara The set of equations \eqref{eq:Arstein-Sontag} is an adaptation of Artstein-Sontag's condition to the  exponential convergence of solutions. Given a control-contraction metric for system \eqref{eq:general system}, Finsler's lemma (cf. \cite[Lemma 11.1]{CalafioreGhaoui2014}) provides stabilizing a feedback law of the form $\delta_u=K\delta_x$ for system \eqref{eq:general system:differential} defined for every $\delta_x\in\mathbb{R}^n$, where $K:\mathbb{R}^n\times\mathbb{R}^m\to\mathbb{R}^{n\times m}$.

@@

\section{Design of Distributed Controllers}\label{sec:Design of Decentralized Controllers}

@@

\mypara The alternating direction method of multipliers (ADMM) provides a framework to compute the block-diagonal matrix $W$ to satisfy the MI~\eqref{eq:MI formulation} in a distributed fashion \cite{BoydParikhChuEtAl2010,MeissenLessardArcakEtAl2015}. To use this methodology, define the set of (resp. sum-of-squares (SOS)) polynomials $\mathbb{R}[x]$ (resp. $\mathbb{SOS}[x]$), and denote the degree of a polynomial $p\in\mathbb{R}[x]$ as $\deg(p)$.

\mypara By applying the change of coordinates $\eta=M\delta$ and defining the matrix $W=M^{-1}$, inequality~\eqref{eq:Artstein-Sontag:inequality} is satisfied, whenever the matrix inequality
\begin{equation}\label{eq:Artstein-Sontag:inequality:dual}
 	-\dot{W}+AW + WA^\top+2\lambda W\prec0
\end{equation} 
holds.

\mypara The search the function $W:\mathbb{R}^n\to\mathbb{R}^{n\times n}$ can be relaxed to a semidefinite program. Whenever $f$ and $B$ are polynomials, this program is cast in terms of the sum-of-squares (SOS) feasibility program \cite{Parrilo2000}. To that end, introduce the vector $v\in\mathbb{R}^n$ and define the polynomials
\begin{align*}
	p_W=&\ v^\top W v\;,\\
	p_{MI}=&\ -v^\top(-\dot{W}+AW + WA^\top+2\lambda W)v\;.
\end{align*}
The feasibility program consists of finding symmetric $Q_W\succeq0$ and $Q_{MI}\succeq0$, called \emph{Gramians}, with appropriate dimensions and such that $p_W=\bar{v}^\top Q_W\bar{v}$ and $p_{MI}=\bar{v}^\top Q_{MI}\bar{v}$ belong to $\mathbb{R}[x,v]$, where $\bar{v}$ is vector of monomials on the variables $x_i,\ldots,x_N$ and $v_1,\ldots,v_N$.

\mypara Once $Q$ has been computed, employing Finsler's lemma \cite[Lemma~11.1]{CalafioreGhaoui2014} pointwisely, the existence of a function $Y:\mathbb{R}^n\to\mathbb{R}^{n\times m}$ satisfying MI~\eqref{eq:MI formulation} is ensured.

\mypara Now that the computation of a positive definite $W$ satisfying inequality~\eqref{eq:Artstein-Sontag:inequality:dual} in terms of SOS polynomials has been explained, the next step is to solve the SOS feasibility using ADMM.

\mypara Define the set $\mathbf{T}=\{Z_1,\ldots,Z_N:Z_W=\mathbin{\mathtt{diag}}(Z_1,\ldots,Z_N)\succeq0,p_{MI}=\bar{v}^\top Z_W\bar{v}\in\mathbb{R}[x,v]\}$. For a set $\mathbf{G}$ define the indicator (also known as barrier) function $H:\{\mathbf{G}\}\times \mathbb{R}^k\to \{0\}\cup\{\infty\}$ as
\begin{equation*}
	H(\mathbf{G},g)=\begin{cases}
									0,&g\in \mathbf{G}\\
									\infty,&\text{Otherwise.}
								\end{cases}
\end{equation*}

\mypara The MI~\eqref{eq:MI formulation} can be cast into the following optimization problem
\begin{equation*}
	\begin{array}{rl}
		\text{minimize}&\displaystyle H(\mathbf{T},(Z_1,\ldots,Z_N))+\sum_{i=1}^N H(\mathbf{S}_{\succ0}^{\alpha_i},Q_{W_i})\\
		\text{subject to}&Q_{W_i}-Z_i=0,\quad\forall i\in\mathbb{N}_{[1,N]}\;,
	\end{array}
\end{equation*}
where $\alpha_i=(2n_i+d_i)!/(d_i!(2n_i)!)$, $d_i=\deg(v_i^\top W_iv_i)$ and $v_i\in\mathbb{R}^{n_i}$.

The ADMM algorithm can be written as
\begin{enumerate}
	\item For each $i\in\mathbb{N}_{[1,N]}$, update the matrix $W_i$ as follows.
	\begin{equation*}
	\begin{array}{rcl}
		Q_{W_i}^{k+1}&=&\mathbin{\mathtt{argmin}}_{Q_{W_i}\in\mathbf{S}_{\succ0}^{\alpha_i}}\left|\left|Q_{W_i}-Z_i^k+\Lambda_i^k\right|\right|_F^2\;,
	\end{array}
\end{equation*}
	where $||\cdot||_F$ stands for the Frobenius norm;

	\item If the matrices $W_1,\ldots,W_N$ belong to $T$, then a solution has been found and the algorithm can be stopped. Otherwise, solve the global problem
	\begin{equation*}
		Z^{k+1}=\mathbin{\mathtt{argmin}}_{Z\in\mathbf{T}}\sum_{i=1}^N\left|\left|Q_{W_i}^{k+1}-Z_i+\Lambda_i^k\right|\right|_F^2\;;
	\end{equation*}
	\item Update $\Lambda$ and return to step 1
	\begin{equation*}
		\Lambda_i^{k+1}=Q_{W_i}^{k+1}-Z_i^{k+1}+\Lambda_i^k\;.
	\end{equation*}
\end{enumerate}
Note that steps 1 and 3 are computed in parellel while step 2 requires to gather all the values of the matrices $Q_{W_i}$ and the values $\Lambda_i$.
